CS6999 Web Data and Text Mining:\\Parallel Text Indexer Project Report

### 2011-06-06 report/report.tex
First draft of literature survey complete
Started final merge

\documentclass[10pt]{report}
\usepackage{graphicx}
\usepackage{setspace}			
\onehalfspacing
% numbering down to subsubsection
\setcounter{secnumdepth}{3}

\title{CS6999 Web Data and Text Mining:\\Parallel Text Indexer Project Report}
\author{Damien DuBios\\Justin Kamerman 3335272\\Raman Singh}
\date{\today}

@@

Algorithm enhancements usually represent the forefront of research in
the field. Work to validate a new concept or algorithm is usually
performed in a very controlled environment and on relatively small
document collections. As the field matures, improvements are harder
fought and often involve increasingly complex algorithms and corpus
approximation methods into which it becomes more difficult to
\textit{fold-in} new documents. In addressing the considerable
computation cost of re-calculating \textit{LSI} \cite{RefWorks:102}
approximations, \cite{RefWorks:101} introduced new update methods
which saw improved average precisions but were forced to conclude that
the improvements were at a computational premium over previous
methods. A semi-discrete matrix decomposition for LSI, \textit{SDD},
performs as well as the original SVD-based LSI method for equal query
times and uses less storage \cite{RefWorks:99}. The decomposition time
of the mew method, however, is significantly larger than the original.

@@

%----------------------------------------
% Motivation
%----------------------------------------
\section{Motivation}
\label{sec:motivation}

@@

%----------------------------------------
% Design Component
%----------------------------------------
\section{Design Component}
\label{sec:designcomponent}
A black-box functional view of the indexing system is show in figure
\ref{fig:blackbox}. The system processes a continuous incoming flow of
documents and distribute them to parallel indexing threads running on
physically separate, heterogeneous nodes.  Document queries are performed
using an evolving search index. The operational balance is to
timeously index new additions to the corpus while servicing concurrent
search requests. A view as to how the components of the indexer system
are deployed is show in figure \ref{fig:deploymentmodel}.

@@

The following subsections detail the implementation of the
architectural layers identified in section \ref{sec:designcomponent}.

@@

@@

%----------------------------------------
% Bibliography
%----------------------------------------
% Change Bibliography title to 'References'
\renewcommand\bibname{References}
\bibliography{bibliography}
\bibliographystyle{IEEEannot}

>>>>

\documentclass[10pt]{report}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{setspace}                   
\onehalfspacing
% numbering down to subsubsection
\setcounter{secnumdepth}{3}

\title{CS6999 Web Data and Text Mining:\\Parallel Text Indexer Project Report}
\author{Damien DuBois\\Justin Kamerman 3335272\\Ramanpreet Singh 3395913}
\date{\today}

@@

Algorithm enhancements usually represent the forefront of research in
the field. Work to validate a new concept or algorithm is usually
performed in a very controlled environment and on relatively small
document collections. As the field matures, improvements are harder
fought and often involve increasingly complex algorithms and corpus
approximation methods into which it becomes more difficult to
\textit{fold-in} new documents. In addressing the considerable
computation cost of re-calculating \textit{LSI} \cite{RefWorks:102}
approximations, \cite{RefWorks:101} introduced new update methods
which saw improved average precisions but were forced to conclude that
the improvements were at a computational premium over previous
methods. A semi-discrete matrix decomposition for LSI, \textit{SDD},
performs as well as the original SVD-based LSI method for equal query
times and uses less storage \cite{RefWorks:99}. The decomposition time
of the new method, however, is significantly larger than the original.

@@

@@

The research design and methods section should include the following:

    an overview of the experimental design

    a detailed description of specific methods to be employed to accomplish the specific aims

    a detailed discussion of the way in which the results will be collected, analyzed, and interpreted

    a projected sequence or timetable (work plan)

    a description of any new methodology used and why it represents an improvement over the existing ones

    a discussion of potential difficulties and limitations and how these will be overcome or mitigated

    expected results, and alternative approaches that will be used if unexpected results are found

    precautions to be exercised with respect to any procedures, situations, or materials that may be hazardous to personnel or human subjects

Suggestion

    Number the sections in this part of the application to correspond to the numbers of the Specific Aims.

    Give sufficient detail. Do not assume that the reviewers will know how you intend to proceed.

    Avoid excessive experimental detail by referring to publications that describe the methods to be employed.

    Publications cited should be by the applicants, if at all possible. Citing someone else's publication establishes that you know what method to use, but citing your own (or that of a collaborator) establishes that the applicant personnel are experienced with the necessary techniques.

    If relevant, explain why one approach or method will be used in preference to others. This establishes that the alternatives were not simply overlooked. Give not only the "how" but the "why."

    If employing a complex technology for the fast time, take extra care to demonstrate familiarity with the experimental details and potential pitfalls. Add a coinvestigator or consultant experienced with the technology, if necessary.

    Document proposed collaborations and offers of materials or reagents of restricted availability with letters from the individuals involved.

\bigskip
The whole idea of the project belongs in being able to reduce the time
to build the index. As our system is building the inverted index by
parsing all the post into the state machine, we can expect that the
complexity will be proportional to the whole number of characters of
all the post we want to build the index for. As a result changing the
number of Posts, the number of characters inside of each post should
have a big impact in our time complexity. Besides, we should also have
a time needed at the very beginning of the running to build the state
machine. We can expect that the building of the state machine needs a
time proportional to the sum of the size of the keywords.

The system proposed in this paper doesn’t take care of various
compression techniques for storing the index. To do that, many
techniques have been explored in the literature. We will just
highlight the different sizes of our indexes according to different
parameters.

Finally we will do a time comparison between our way of building the
inverted index and the naive way.  All the time complexities tests are
done on a single computer and it is important to understand that all
these processes can be parallelized. A future work would be to install
the system on multiple computers and explore the gain in time by
running the code simultaneously on these computers.

%----------------------------------------
% Design
%----------------------------------------
\section{Design}
\label{sec:design}
A black-box functional view of the indexing system is show in figure
\ref{fig:blackbox}. The system processes a continuous incoming flow of
documents and distribute them to parallel indexing threads running on
physically separate, heterogeneous nodes.  Document queries are performed
using an evolving search index. The operational balance is to
timeously index new additions to the corpus while servicing concurrent
search requests. A view as to how the components of the indexer system
are deployed is show in figure \ref{fig:deploymentmodel}.

@@

The following subsections detail the implementation of the
architectural layers identified in section \ref{sec:design}.

@@

%----------------------------------------
% Number of Keywords
%----------------------------------------
\subsection{Number of Keywords}
To be able to study the impact of different number of keywords, we
built a keywords database with 10.000 most popular English
words. These 10.000 words come from:
http://www.mit.edu/~ecprice/wordlist.10000 All the experiments in this
part are done with 50.000 posts randomly downloaded from internet. The
idea is to get some heterogeneous length and type of post such that it
is realistic.  The experiment shown here consist on building the index
for different number of keywords from the 10.000 keywords database and
explore the time and space complexity for the building of the state
machine and the building of the index once the state machine is built.

\subsubsection{Time Complexity}
\begin{itemize}
\item \textbf{State Machine Construction:}
  The time taken to construct the Aho-Corasick state machine was
  measured for different numbers of keywords. The results of this test
  is shown in figure \ref{fig:numkeywordstimecomplexbuildsm} and construction time
  can easily be considered linear with respect to the number of
  keywords. This is consistent with \cite{RefWorks:103} which proves
  that the state machine construction algorithm is linearly
  proportional to the sum of the lengths of the keywords used to
  construct the state machine.

  \begin{figure}
    \begin{center}
      \includegraphics[width=\textwidth,height=!]{numkeywordstimecomplexbuildsm}
    \end{center}
    \caption{Time to build Aho-Corasick state machine as a function of
      the number of keywords. Time taken is shown in miliseconds on the
      y-axis and the number of keywords on the x-axis}
    \label{fig:numkeywordstimecomplexbuildsm}
  \end{figure}

\item \textbf{Index Construction:} 
  The time taken for a state machines to build an inverted index of our
  corpus was measured. The test was repeated using state machines
  constructed with a varying number of keywords and the results plotted
  in figure \ref{fig:numkeywordstimecomplexbuildindex}. According to
  \cite{RefWorks:103} which proves that the number of state transitions
  involved in processing an input string is independent of the number of
  keywords used to construct the state machine, we would not expect the
  time complexity to increase with the size of the keyword set. However,
  as can be seen in \ref{fig:numkeywordstimecomplexbuildindex}, the time
  to build the index increases logarithmically with the number of
  keywords. This is likely due to the fac that as the number of keywords
  increases, so does the size of the index and the number of database
  interacts required to store and update the index.

  \begin{figure}
    \begin{center}
      \includegraphics[width=\textwidth,height=!]{numkeywordstimecomplexbuildindex}
    \end{center}
    \caption{Time to build the inverted index as a function of the
      number of keywords. Time taken is shown in miliseconds on the
      y-axis and the number of keywords on the x-axis}
    \label{fig:numkeywordstimecomplexbuildindex}
  \end{figure}

\end{itemize}

\subsubsection{Space Complexity}
As we would have expected it, the space used by the index in the
database is proportional to the number of words we are using. It also
means that our words have the same probability at the beginning or at
the end of our 12,000 words dictionnary. This result is reflected in
figure \ref{fig:numkeywordsspacecomplexbuildindex}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{numkeywordsspacecomplexbuildindex}
  \end{center}
    \caption{Size of index as a function of the number of
      keywords. The number of database rows used to store the index is
      shown on the y-axis and the number of keywords on the x-axis} 
    \label{fig:numkeywordsspacecomplexbuildindex}
\end{figure}

%----------------------------------------
% Size of Corpus
%----------------------------------------
\subsection{Size of Corpus}
To be able to study the impact of different number of Posts, we built
a posts database with 400.000 posts randomly downloaded from
internet. Then the keywords used to build the index are simply the
first 5.000 from the previous keywords database.

The experiment shown here consist on building the index for different
number of Posts and explore the time and space complexity for the
building of the state machine and the building of the index once the
state machine is built. We can expect that the time to build the State
machine won’t change with the number of posts and the time to build
the index will increase almost linearly with the number of posts.

\subsubsection{Time Complexity}
As we were expecting, the time to build the index with the state
machine strategy is linear with respect to the number of post, as
depicted in figure \ref{fig:corpsizetimecomplexbuildindex}. However, to be able
to perform this experiment we had to run the indexer program with
batches of  10,000 posts due to Java heap space problems. As a result, the state
machine is rebuilt every 10,000 posts.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{corpsizetimecomplexbuildindex}
  \end{center}
    \caption{Time to build the inverted index as a function of the
      size of the corpus. Time taken is shown in miliseconds on the
      y-axis and the number of documents in the collection on the
      x-axis} 
    \label{fig:corpsizetimecomplexbuildindex}
\end{figure}

\subsubsection{Space Complexity}
As expected, the space used to sptore the index in the database is
proportional to the size of the corpus. This result is shown in figure
\ref{fig:corpsizespacecomplexbuildindex}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{corpsizespacecomplexbuildindex}
  \end{center}
    \caption{Size of index as a function of the size of the
      corpus. The number of database rows used to store the index is
      shown on the y-axis and the number of documents in the
      collection on the x-axis} 
    \label{fig:corpsizespacecomplexbuildindex}
\end{figure}

%----------------------------------------
% Document Size
%----------------------------------------
\subsection{Document Size} 
As we have seen before, the time to build the index is almost
proportional with the number of post. Theoretically this time should
be proportional to the total length of all posts. Let’s focus on this
aspect and study the impact of the length of the post on the time to
build the state machine. To do that we built a 5.000 posts database
with posts that are all composed of 20.000 characters. Then, we ran
our system on different length of substring of these posts.

\subsubsection{Time Complexity}
\begin{itemize}
\item \textbf{State Machine Construction:} 
  As expected, state machine construction time
  does not depend on the length of the documents. Nonetheless it is
  interesting to compare figures \ref{fig:docsizetimecomplexbuildsm}
  and \ref{fig:docsizetimecomplexbuildindex}, showing the construction
  times for the state machine and inverted index relative to the
  average document size. The comparison shows that state machine
  construction time is several orders of magnitude smaller than index
  construction and negligible by comnparison. Our state machine is
  built in 770 ms on average for 1000 keywords with a standard
  deviation of 18ms.

  \begin{figure}
    \begin{center}
      \includegraphics[width=\textwidth,height=!]{docsizetimecomplexbuildsm}
    \end{center}
    \caption{Time to build Aho-Corasick state machine as a function of
      the average document size. Time taken is shown in miliseconds on the
      y-axis and the average document size in characters on the x-axis}
    \label{fig:docsizetimecomplexbuildsm}
  \end{figure}

\item \textbf{Index Construction:} 
  As expected the time to build the indexer is almost linear with
  respect to the average document size. This relationship is shown in
  figure \ref{fig:docsizetimecomplexbuildindex}.

  \begin{figure}
    \begin{center}
      \includegraphics[width=\textwidth,height=!]{docsizetimecomplexbuildindex}
    \end{center}
    \caption{Time to build an inverted index as a function of the
      average document size. Time taken is shown in miliseconds on the
      y-axis and the average document size in characters on the x-axis} 
    \label{fig:docsizetimecomplexbuildindex}
  \end{figure}

\end{itemize}

\subsubsection{Space Complexity}
Here we are studying the number of records in our index database as a
function of the average document size. As expected, the larger the documents
the more records created to represent the index. However, this is not linear
since if a word occurs many times in the same post we just save it
once in the database. This relationship is shown in figure
\ref{fig:docsizespacecomplexbuildindex}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{docsizespacecomplexbuildindex}
  \end{center}
  \caption{Size of index as a function of the average document
    size. The number of database rows used to store the index is
      shown on the y-axis and the average document size in characters
      on the x-axis}  
  \label{fig:docsizespacecomplexbuildindex}
\end{figure}

%----------------------------------------
% Comparison with Naive Index Builder
%----------------------------------------
\subsection{Comparison with Na\"{a}ve Index Builder}
This part focuses on the improvement that our system with respect to
time complexity in comparison with a na\"{i}e index builder. The na\"{i}ve
indexer builder has the same input as our system i.e. a list of
keywords and a list of documents, and the same output i.e. an inverted
index corresponding to the keywords and documents. Algorithm
\ref{alg:invertedindex} is used in principal for both implementations.

\begin{algorithm}
\caption{Build inverted index}
\label{alg:invertedindex}
\begin{algorithmic}
  \FORALL{$keyword$}
  \STATE Build a vector $v$
  \FORALL {$document$}
  \IF {$keyword$ \epsilon \; $document$}
  \STATE Add the row (keyword.id, document.id) into $v$
  \ENDFOR
  \STATE Save $v$ into the database
  \ENDFOR
\end{algorithmic}
\end{algorithm}

As we can see this algorithm has a complexity that directly depends on
the number of keywords n, the number of document m and the length of
document since we have to check in each document for each
keyword. Moreover, we decided to limit as much as possible the
accesses to the database since it is really time consuming in
Java. That’s why we do one access to the database for each keywords
and not for each document.

Note: In the state machine we do as much accesses to the database as
the number of document (in the worst case meaning we find at least one
keyword per document).

Let’s compare the results for this algorithm with our system results
for different number of keywords and posts.

\subsubsection{Analytical}

\subsubsection{Experimental}

@@

The experimental results tend to prove what we could expect from
theory. Our system is faster than the naive indexer whatever
parameters we are changing. In particular when the number of document
is becoming really important, the State machine approach prove to be
very efficient. Moreover, as soon as we are dealing with a number of
keywords that can be representative of industrial needs, the state
machine prove also to be far faster than the naive indexer.

Future work on this strategy would be to compare this approach with
less naive strategies to build an index. Moreover, as said in the
introduction, it would probably be interesting to add compression
methods to our algorithm in order to reduce the index size in the
database.

%----------------------------------------
% Bibliography
%----------------------------------------
% changes default name Bibliography to References
\renewcommand\bibname{References}
\bibliography{bibliography}
\bibliographystyle{IEEEannot}
